---
title: "GnomAD exomes vs genomes: ClusterBurden 'null' model"
author: "Adam Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GnomAD exomes vs genomes: ClusterBurden 'null' model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7
)
```


# BIN-test no coverage control  

1. Generate GnomAD exomes 'controls'
2. Generate GnomAD genomes 'cases'
3. Switch affection status to GnomAD genomes 'cases' to 1 (aff=1)
4. Remove switch to turn on automatic coverage 
5. Calculate BIN-test p-values with no coverage control

```{r}
library(ClusterBurden)

controls = collect_gnomad_controls(dataset="exome")
cases = collect_gnomad_controls(dataset="genome")
cases[,aff:=1]

attributes(controls)$provenance = attributes(cases)$provenance = NULL

binpval = BIN_test_WES(cases, controls)

head(binpval)


```

## Proportion of significant results 

```{r}
# nominal significance 
binpval[,sum(BIN.test_pvalue < 0.05, na.rm=T)/.N]

# Bonferonni significance 
binpval[,sum(BIN.test_pvalue < 0.05/.N)/.N]
```

## Lambda (inflation)

```{r}
# function to generate lambda inflation metric
lambda = function(p) median(qchisq(p, df=1, lower.tail=FALSE)) / qchisq(0.5, 1)

# lambda across all non-NA p-values
lambda(binpval[!is.na(BIN.test_pvalue), BIN.test_pvalue])

# lambda across p-values not in giant proteins i.e. > 2000 residues
lambda(binpval[!is.na(BIN.test_pvalue) & !grepl("high", pl_flag), BIN.test_pvalue])
```

## Manhattan plot

```{r}
manhattan(binpval, "bin-test", 10, SCALE=0.5)
```

# BIN-test with coverage control

1. Turn back on the switch for automatic coverage control
5. Calculate BIN-test p-values with automatic coverage control

```{r}
attributes(controls)$provenance = attributes(cases)$provenance = "auto"

binpval = BIN_test_WES(cases, controls)

binpval
```

## Significant results 

It is clear that coverage differences was the main driver of significant results. Lambda is now reduced to become below 1 however nominally significant results do exceed 5%. The manhattan plot reveals some Bonferonni significant results indicating either some uncontrolled for confounders (e.g. ancestry) or true differences between the exomes and genomes cohorts (unlikely but not impossible). 

```{r}
# nominal significance 
binpval[,sum(BIN.test_pvalue < 0.05, na.rm=T)/.N]

# Bonferonni significance 
binpval[,sum(BIN.test_pvalue < 0.05/.N, na.rm=T)/.N]

# function to generate lambda inflation metric
lambda = function(p) median(qchisq(p, df=1, lower.tail=FALSE)) / qchisq(0.5, 1)

# lambda across all non-NA p-values
lambda(binpval[!is.na(BIN.test_pvalue), BIN.test_pvalue])

# manhattan
manhattan(binpval, "bin-test", 10, SCALE=0.5)
```

Alongside the p-values there are flags to alert the user to potential confounders including pl_flag for protein length and cov_flag1 and cov_flag2 for coverage. After removing genes with coverage concern level flagged as high (or very high) the same calculations are made.

```{r}
binpval = binpval[!grepl("high", cov_flag1) & !grepl("high", cov_flag2) & !grepl("high", pl_flag)]

# nominal significance 
binpval[,sum(BIN.test_pvalue < 0.05, na.rm=T)/.N]

# Bonferonni significance 
binpval[,sum(BIN.test_pvalue < 0.05/.N, na.rm=T)/.N]

# function to generate lambda inflation metric
lambda = function(p) median(qchisq(p, df=1, lower.tail=FALSE)) / qchisq(0.5, 1)

# lambda across all non-NA p-values
lambda(binpval[!is.na(BIN.test_pvalue), BIN.test_pvalue])

# manhattan
manhattan(binpval, "bin-test", 10, SCALE=0.5)
```
